### 2017-05-22 doc/categories.tex

\begin{figure*}[t]
  \centering
  \hspace*{-0.2in}
  \setlength{\tabcolsep}{0pt}
  \begin{tabular}{cc}
  \includegraphics[scale=0.55]{imaginet-omission-ratio-pos-boxplot.png} &
  \includegraphics[scale=0.55]{imaginet-omission-ratio-dep-boxplot.png} \\  
  \end{tabular}
  \caption{Distributions of log ratios of omission scores of {\sc Textual} to {\sc Visual} per
    POS (left) and deprel labels (right). Only labels which occur at least 500 times are included.}
\label{fig:omission-imaginet-ratio}
\end{figure*}

\begin{figure*}[t]
  \centering
  \hspace*{-0.2in}
  \setlength{\tabcolsep}{0pt}
  \begin{tabular}{cc}
  \includegraphics[scale=0.55]{imaginet-omission-quotient-pos-boxplot.png} &
  \includegraphics[scale=0.55]{imaginet-omission-quotient-dep-boxplot.png} \\  
  \end{tabular}
  \caption{Distributions of log ratios of omission scores of {\sc LM} to {\sc Textual} per
    POS (left) and deprel labels (right). Only labels which occur at least 500 times are included.}
\label{fig:omission-imaginet-ratio}
\end{figure*}

@@

We fit four linear models which predict the omission 
scores per token with the following predictor variables:

\begin{enumerate}
	\item {\sc LM word}: word type
	\item {\sc LM deprel}: word type, dependency label and their interaction 
	\item {\sc LM position}: word type, position (binned as first, second, third, middle,
	antepenult, penult, last) and their interaction
	\item {\sc LM full}: word type, dependency label, position, word-deprel interaction, 
	word-position interaction
\end{enumerate}

\noindent We split the whole MSCOCO validation set into two parts, fit the 
models (regularized via L2 penalty) on the first part, and compute the 
proportion of variance explained on the second part. 
For comparison we also train a version of {\sc Imaginet},
where the GRU in the {\sc Visual} pathway is replaced by a simple vector-summation operation
over word-embeddings and thus does not have access to sequential cues. This
model we refer to as {\sc Sum}.

Figure~\ref{fig:rsquared} shows the increase in $R^2$  for the models 
per pathway relative to {\sc LM word}. The plot reveals that 
adding additional information on top of lexical features in case of {\sc Sum} 
does increase the explained variance slightly, which is probably due to 
the unseen words in the held out set. More interestingly, the
relative increase in $R^2$ between {\sc LM word} and {\sc LM full} in
case of {\sc Visual} is larger, suggesting that 
the omission scores in case of {\sc Visual} do not only depend on the words themselves,
but their grammatical function and position in sentences.  The raw  $R^2$
scores show that for the {\sc Textual} model the word-type predicts the omission-score
to much smaller degree compared to {\sc Visual} (0.38 against 0.48), and the 
increase in $R^2$ shows that the position of the word adds considerable 
amount of information.

\begin{figure}
\centering
  \includegraphics[scale=0.35]{omission-stat/position-new.png}
\caption{Proportion of variance in omission scores explained by the models
 for {\sc Sum}, {\sc Visual} and {\sc Textual} pathways.}
\label{fig:rsquared}
\end{figure}

>>>>

\begin{figure*}[t]
  \centering
  \hspace*{-0.2in}
  \setlength{\tabcolsep}{0pt}
  \begin{tabular}{cc}
  \includegraphics[scale=0.55]{imaginet-omission-ratio-pos-boxplot.png} &
  \includegraphics[scale=0.55]{imaginet-omission-ratio-dep-boxplot.png} \\  
  \end{tabular}
  \caption{Distributions of log ratios of omission scores of {\sc Textual} to {\sc Visual} per
    POS (left) and dependency labels (right). Only labels which occur at least 500 times are included.}
\label{fig:omission-imaginet-ratio}
\end{figure*}

\begin{figure*}[t]
  \centering
  \hspace*{-0.2in}
  \setlength{\tabcolsep}{0pt}
  \begin{tabular}{cc}
  \includegraphics[scale=0.55]{imaginet-omission-quotient-pos-boxplot.png} &
  \includegraphics[scale=0.55]{imaginet-omission-quotient-dep-boxplot.png} \\  
  \end{tabular}
  \caption{Distributions of log ratios of omission scores of {\sc LM} to {\sc Textual} per
    POS (left) and dependency labels (right). Only labels which occur at least 500 times are included.}
\label{fig:omission-imaginet-ratio}
\end{figure*}

@@

We fit four L2-penalized linear regression models which predict the omission 
scores per token with the following predictor variables:

\begin{enumerate}
	\item {\sc LR word}: word type
	\item {\sc LR deprel}: word type, dependency label and their interaction 
	\item {\sc LR position}: word type, position (binned as first, second, third, middle,
	antepenult, penult, last) and their interaction
	\item {\sc LR full}: word type, dependency label, position, word:dependency interaction, 
	word:position interaction
\end{enumerate}

\noindent We split the whole MSCOCO validation set into two parts, fit the 
models (regularized via L2 penalty) on the first part, and compute the 
proportion of variance explained on the second part. 
For comparison we also train a version of {\sc Imaginet},
where the GRU in the {\sc Visual} pathway is replaced by a simple vector-summation operation
over word-embeddings and thus does not have access to sequential cues. This
mposodel we refer to as {\sc Sum}.

Figure~\ref{fig:rsquared} shows the increase in $R^2$  for the models 
per pathway relative to {\sc LR word}. The plot reveals that 
adding additional information on top of lexical features in case of {\sc Sum} 
does increase the explained variance slightly, which is probably due to 
the unseen words in the held out set. More interestingly, the
relative increase in $R^2$ between {\sc LR word} and {\sc LR full} in
case of {\sc Visual} is larger, suggesting that 
the omission scores in case of {\sc Visual} do not only depend on the words themselves,
but their grammatical function and position in sentences.  The raw  $R^2$
scores show that for the {\sc Textual} model the word-type predicts the omission-score
to much smaller degree compared to {\sc Visual} (0.38 against 0.48), and the 
increase in $R^2$ shows that the position of the word adds considerable 
amount of information.

\begin{figure}
\centering
  \includegraphics[scale=0.35]{omission-stat/position-new.png}
\caption{Proportion of variance in omission scores explained by the
  linear regression models
 for {\sc Sum}, {\sc Visual} and {\sc Textual} pathways, relative to
 regressing on word identity only. }
\label{fig:rsquared}
\end{figure}
